---
title: "Assignment 1"
author: "Inn Kyung Seo"
output:
  html_document: default
  pdf_document: default
---
## Purpose
The purpose of this assignment is to set up and use R and GitHub for data analysis.  
For this task, I downloaded the Iris dataset from the UCI Machine Learning Repository, imported it into R, generated descriptive statistics, created a new variable, and produced visualizations.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

iris <- read.csv(file.choose(), header = FALSE) 
# Data source: UCI Machine Learning Repository – Iris Dataset
iris <- read.csv(
  "https://archive.ics.uci.edu/ml/machine-learning-databases/iris/iris.data",
  header = FALSE
)

```{r}
colnames(iris) <- c("Sepal.Length","Sepal.Width","Petal.Length","Petal.Width","Species")


head(iris, 3)

summary(iris)


table(iris$Species)
```

```{r}
iris$Petal.Length.Sq <- iris$Petal.Length^2
head(iris, 3)

hist(iris$Sepal.Length,
     main = "Histogram of Sepal Length",
     xlab = "Sepal Length",
     col = "lightblue")

plot(iris$Petal.Length, iris$Petal.Width,
     col = as.factor(iris$Species),
     main = "Petal Length vs Width by Species",
     xlab = "Petal Length", ylab = "Petal Width",
     pch = 17)
legend("topleft", legend = levels(as.factor(iris$Species)),
       col = 1:3, pch = 17)
```
## Data Source
Data Source: UCI Machine Learning Repository – Iris Dataset  
https://archive.ics.uci.edu/ml/machine-learning-databases/iris/iris.data


